---
title: "Data 624 - Project 1"
author: "Baron Curtin"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    toc: true
    theme: cayman
    highlight: github
    df_print: paged
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
knitr::opts_chunk$set(warning = FALSE)
libs <- c("tidyverse", "knitr", "kableExtra", "fpp2", "forecast", "imputeTS", "xts", "readxl", "xlsx", "glue", "gridExtra", "lubridate")
loadPkg <- function(x) {
  if(!require(x, character.only = T)) install.packages(x, dependencies = T, repos = "http://cran.us.r-project.org")
  require(x, character.only = T)
}
lapply(libs, loadPkg)
```

```{r, include=FALSE}
# data import S01, S03 (No V02), S04
csvs <- c("S01.csv", "S03.csv", "S04.csv")
grpVars <- c("Var01", "Var02", "Var05", "Var07")

datasets <- csvs %>%
  map(function(csv) {
    data <- read_csv(csv) %>%
      mutate(SeriesInd = as.Date(SeriesInd, origin ="1899-12-30"),
             LubDate = ymd(SeriesInd),
             DayOfWeek = wday(LubDate, label = T)) %>%
      slice(1:(n() - 140)) %>%
      select(-group)
  }) %>%
  setNames(csvs)
```

## Exploratory Analysis
```{r}
datasets %>% iwalk(function(data, csv) {
  print(csv)
  print(summary(data))
})
```

Using the summary function:
* S01 has 2 total NAs in Var01
* S03 has 8 total NAs, 4 in each of Var05 and Var07
* S04 has 2 total NAs in Var02
* There are no Wednesdays or Thurdays in the dataset
* We will have to impute these NAs prior to forecasting

### Plots
```{r, include=FALSE}
# function to generate plots
genPlots <- function(data, x, y, title) {
  # generate plot
  g <- ggplot(data, aes_(as.name(x), as.name(y))) +
    geom_line() +
    ggtitle(title)
  
  return(g)
}

genBoxPlots <- function(data, var) {
  # reformat data for easy plotting
  data <- data %>%
    select(SeriesInd, DayOfWeek, var) %>%
    gather("var", "amount", -SeriesInd, -DayOfWeek)
  
  # generate plot
  g <- ggplot(data, aes_(as.name("var"), as.name("amount"))) +
    geom_boxplot()
  return(g)
}

arrangePlots <- function(p1, p2, rows = 2) {
  grid.arrange(p1, p2, nrow = rows)
}

x <- "SeriesInd"
```

S01
```{r}
g <- "S01.csv"

plots <- c('Var01', 'Var02') %>%
  map(~ genPlots(datasets[[g]], "SeriesInd", .x, g))

arrangePlots(plots[[1]], plots[[2]])
```

The plots reveal:
* Var02 contains a really large outlier
* This will have to be dealt with prior to forecasting


```{r}
plots <- c('Var01', 'Var02') %>%
  map(~ genBoxPlots(datasets[[g]], .x))

arrangePlots(plots[[1]], plots[[2]], 1)
```

The boxplots reveal:
* Var02 contains a good amount of outliers
* This will have to be dealt with prior to forecasting


S03
```{r}
g <- "S03.csv"

plots <- c('Var05', 'Var07') %>%
  map(~ genPlots(datasets[[g]], "SeriesInd", .x, g))

arrangePlots(plots[[1]], plots[[2]])
```

The plots reveal:
* Var05 and Var07 are relatively similar


```{r}
plots <- c('Var05', 'Var07') %>%
  map(~ genBoxPlots(datasets[[g]], .x))

arrangePlots(plots[[1]], plots[[2]], 1)
```


S04
```{r}
g <- "S04.csv"

plots <- c('Var01', 'Var02') %>%
  map(~ genPlots(datasets[[g]], "SeriesInd", .x, g))

arrangePlots(plots[[1]], plots[[2]])
```

The plots reveal:
* Var02 contains a really large outlier
* This will have to be dealt with prior to forecasting


```{r}
plots <- c('Var01', 'Var02') %>%
  map(~ genBoxPlots(datasets[[g]], .x))

arrangePlots(plots[[1]], plots[[2]], 1)
```


## Data Preprocessing
```{r, results='hide'}
# identify and remove outliers
identifyOutliers <- function(data, var) {
  if(var %in% colnames(data)) {
    outliers <- boxplot(data[var])$out
    return(outliers)
  }
}

replaceOutliers <- function(data, var, outliers) {
  if(var %in% colnames(data)) {
    data[which(data[var] %in% outliers), var] <- NA
    return(data)
  }
}

processedData <- datasets %>%
  map(function(data) {
    df <- data
    
    for (var in grpVars) {
      if(var %in% colnames(df)) {
        outliers <- identifyOutliers(df, var)
        df <- replaceOutliers(df, var, outliers)
      }
    }
    return(df)
  })
```



```{r}
# impute missing values
imputeNa <- function(data, var) {
  if(var %in% colnames(data)) {
    imp <- na.kalman(data[[var]], model = "auto.arima")
    return(imp)
  }
}

processedData <- processedData %>%
  map(function(data) {
    for (var in grpVars) {
      if (var %in% colnames(data)) {
        data[[var]] <- imputeNa(data, var)
      }
    } 
    return(data)
  })
```


Create the Timeseries Objects
```{r}
# create timeseries
createTs <- function(data, var) {
  if(var %in% colnames(data)) {
    timeSer <- xts(data[[var]], order.by = data[["SeriesInd"]])
    return(timeSer)
  }
}

timeSeries_ <- processedData %>%
  map(function(data) {
    vs <- grpVars[grpVars %in% colnames(data)]
    tsList <- vs %>%
      map(function(var) {
        if (var %in% colnames(data)) {
          myTs <- xts(data[[var]], order.by = data[["SeriesInd"]])
          return(myTs)
        }
      })
    return(tsList)
  })
```


Box-Cox Transform
```{r}
bxcxTransform <- function(data) {
  lmda <- BoxCox.lambda(data)
  data <- BoxCox(data, lmda)
  return(data)
}

timeSeries_ <- timeSeries_ %>%
  map(function(tsList) {
    tsList <- tsList %>%
      map(~ bxcxTransform(.x))
    return(tsList)
  })
```


### Modelling
```{r}
models <- timeSeries_ %>%
  map(function(tsList) {
    tsList <- tsList %>%
      map(~ auto.arima(.x, stepwise=FALSE, approximation=FALSE))
  })

```


### Forecasting
```{r}
forecasts <- models %>%
  map(function(modelList) {
    modelList <- modelList %>%
      map(function(model) {
        results <- list(
          checkResiduals = checkresiduals(model),
          forecast = forecast(model, h = 140)
        )
        return(results)
      })
    return(modelList)
  })
```

